# Data Visualisation {#ggplot}

Take the [quiz](#ggplot-quiz) to see if you need to review this chapter.

## Learning Objectives

### Basic

1. Understand what types of graphs are best for [different types of data](#vartypes)
    + 1 discrete
    + 1 continuous
    + 2 discrete
    + 2 continuous
    + 1 discrete, 1 continuous
    + 3 continuous
2. Create common types of graphs with ggplot2
    + [`geom_bar()`](#geom_bar)
    + [`geom_density()`](#geom_density)
    + [`geom_freqpoly()`](#geom_freqpoly)
    + [`geom_histogram()`](#geom_histogram)
    + [`geom_violin()`](#geom_violin)
    + [`geom_boxplot()`](#geom_boxplot)
    + [`geom_col()`](#geom_col)
    + [`geom_point()`](#geom_point)
    + [`geom_smooth()`](#geom_smooth)
3. Set custom labels
4. Represent factorial designs with different colours or facets
5. [Save plots](#ggsave) as an image file
    
### Intermediate

6. Superimpose different types of graphs
7. Add lines to graphs
8. Deal with [overlapping data](#overlap)
9. Create less common types of graphs
    + [`geom_tile()`](#geom_tile)
    + [`geom_density2d()`](#geom_density2d)
    + [`geom_bin2d()`](#geom_bin2d)
    + [`geom_hex()`](#geom_hex)
    + [`geom_count()`](#geom_count)
10. Use the [`viridis`](#viridis) package to set colours

### Advanced

11. Arrange plots in a grid using [`cowplot`](#cowplot)
12. Adjust axes (e.g., flip coordinates, set axis limits)
13. Change the [theme](#theme)
14. Create interactive graphs with [`plotly`](#plotly)


## Resources

* [Look at Data](http://socviz.co/look-at-data.html) from [Data Vizualization for Social Science](http://socviz.co/)
* [Chapter 3: Data Visualisation](http://r4ds.had.co.nz/data-visualisation.html) of *R for Data Science*
* [Chapter 28: Graphics for communication](http://r4ds.had.co.nz/graphics-for-communication.html) of *R for Data Science*
* [ggplot2 cheat sheet](https://www.rstudio.com/wp-content/uploads/2016/11/ggplot2-cheatsheet-2.1.pdf)
* [ggplot2 documentation](http://ggplot2.tidyverse.org/reference/)
* [The R Graph Gallery](http://www.r-graph-gallery.com/) (this is really useful)
* [Top 50 ggplot2 Visualizations](http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html)
* [R Graphics Cookbook](http://www.cookbook-r.com/Graphs/) by Winston Chang
* [The viridis color palettes](https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html)
* [ggplot extensions](https://www.ggplot2-exts.org/)
* [plotly](https://plot.ly/ggplot2/) for creating interactive graphs


## Setup

```{r setup, results = 'hide', warning = FALSE, message = FALSE}

# libraries needed for these graphs
library(tidyverse)
library(viridis)
library(plotly)

# cowplot will change the default theme of graphs, so we're loading it later
# library(cowplot) 

```

## Common Variable Combinations {#vartypes}

* 1 discrete
* 1 continuous
* 2 discrete
* 2 continuous
* 1 discrete, 1 continuous
* 3 continuous

```{block, type="try"}
Before you read ahead, come up with an example of each type of variable combination and sketch the types of graphs that would best display these data.
```

## Data

Here we've created some data frames with different types of data. 

* `pets` has a column with pet type
* `pet_happy` has `happiness` and `age` for 500 dog owners and 500 cat owners
* `x_vs_y` has two correlated continuous variables (`x` and `y`)
* `overlap` has two correlated ordinal variables and 1000 observations so there is a lot of overlap
* `overplot` has two correlated continuous variables and 10000 observations

```{r set-seed, echo = FALSE}
set.seed(8675309)
```

```{r sim-data}

pets <- tibble(
  pet = sample(
    c("dog", "cat", "ferret", "bird", "fish"), 
    100, 
    TRUE, 
    c(0.45, 0.40, 0.05, 0.05, 0.05)
  )
)

pet_happy <- tibble(
  pet = rep(c("dog", "cat"), each = 500),
  happiness = c(rnorm(500, 55, 10), rnorm(500, 45, 10)),
  age = rpois(1000, 3) + 20
)

x_vs_y <- tibble(
  x = rnorm(100),
  y = x + rnorm(100, 0, 0.5)
)

overlap <- tibble(
  x = rbinom(1000, 10, 0.5),
  y = x + rbinom(1000, 20, 0.5)
)

overplot <- tibble(
  x = rnorm(10000),
  y = x + rnorm(10000, 0, 0.5)
)

```

```{block, type"try"}
First, think about what kinds of graphs are best for representing these different types of data.
```

## Basic Plots

### Bar plot {#geom_bar}

Bar plots are good for categorical data where you want to represent the count.

```{r barplot, fig.cap="Bar plot"}
ggplot(pets, aes(pet)) +
  geom_bar()
```

### Density plot
<a name="geom_density"></a>
Density plots are good for one continuous variable, but only if you have a fairly 
large number of observations.

```{r density, fig.cap="Density plot"}

ggplot(pet_happy, aes(happiness)) +
  geom_density()

```

You can represent subsets of a variable by assigning the category variable to 
the argument `group`, `fill`, or `color`. 

```{r density-grouped, fig.cap="Grouped density plot"}

ggplot(pet_happy, aes(happiness, fill = pet)) +
  geom_density(alpha = 0.5)

```

```{block, type="try"}
Try changing the `alpha` argument to figure out what it does.
```

### Frequency Polygons {#geom_freqpoly}

If you don't want smoothed distributions, try `geom_freqpoly()`.

```{r freqpoly, fig.cap="Frequency ploygon plot"}

ggplot(pet_happy, aes(happiness, color = pet)) +
  geom_freqpoly(binwidth = 1)
  
```

``````{block, type="try"}
Try changing the `binwidth` argument to 5 and 0.1. How do you figure out the right value?
```

### Histogram {#geom_histogram}

Histograms are also good for one continuous variable, and work well if you don't have many observations. Set the `binwidth` to control how wide each bar is.

```{r histogram, fig.cap="Histogram"}

ggplot(pet_happy, aes(happiness)) +
  geom_histogram(binwidth = 1, fill = "white", color = "black")
  
```

```{block, type="info"}
Histograms in ggplot look pretty bad unless you set the `fill` and `color`.
```

If you show grouped histograms, you also probably want to change the default `position` argument.

```{r histogram-grouped, fig.cap="Grouped Histogram"}

ggplot(pet_happy, aes(happiness, fill=pet)) +
  geom_histogram(binwidth = 1, alpha = 0.5, position = "dodge")
  
```

```{block, type="try"}
Try changing the `position` argument to "identity", "fill", "dodge", or "stack".
```

### Column plot {#geom_col}

Column plots are the worst way to represent grouped continuous data, but also one of the most common.

To make column plots with error bars, you first need to calculate the means, error bar uper limits (`ymax`) and error bar lower limits (`ymin`) for each category. You'll learn more about how to use the code below in the next two lessons.

```{r colplot, fig.cap="Column plot"}

# calculate mean and SD for each pet
pet_happy %>%
  group_by(pet) %>%
  summarise(
    mean = mean(happiness),
    sd = sd(happiness)
  ) %>%
ggplot(aes(pet, mean, fill=pet)) +
  geom_col(alpha = 0.5) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.25) +
  geom_hline(yintercept = 40)

```

```{block, type="try"}
What do you think `geom_hline()` does?
```

### Boxplot {#geom_boxplot}

Boxplots are great for representing the distribution of grouped continuous 
variables. They fix most of the problems with using barplots for continuous data.

```{r boxplot, fig.cap="Box plot"}

ggplot(pet_happy, aes(pet, happiness, fill=pet)) +
  geom_boxplot(alpha = 0.5)

```

### Violin plot {#geom_violin}

Violin pots are like sideways, mirrored density plots. They give even more information than a boxplot about distribution and are especially useful when you have non-normal distributions.

```{r violin, fig.cap="Violin plot"}

ggplot(pet_happy, aes(pet, happiness, fill=pet)) +
  geom_violin(
    trim = FALSE,
    draw_quantiles = c(0.25, 0.5, 0.75), 
    alpha = 0.5
  )

```

```{block, type="try"}
Try changing the numbers in the `draw_quantiles` argument.
```


### Scatter plot {#geom_point}

Scatter plots are a good way to represent the relationship between two continuous variables.

```{r scatter, fig.cap="Scatter plot using geom_point()"}

ggplot(x_vs_y, aes(x, y)) +
  geom_point()

```

### Line graph {#geom_smooth}

You often want to represent the relationship as a single line.

```{r line, fig.cap="Line plot using geom_smooth()"}
ggplot(x_vs_y, aes(x, y)) +
  geom_smooth(method="lm")
```
 
 
## Save as File
<a name="ggsave"></a>
You can save a ggplot using `ggsave()`. It saves the last ggplot you made, 
by default, but you can specify which plot you want to save if you assigned that 
plot to a variable.

You can set the `width` and `height` of your plot. The default units are inches, 
but you can change the `units` argument to "in", "cm", or "mm".


```{r ggsave-example, eval = FALSE}

box <- ggplot(pet_happy, aes(pet, happiness, fill=pet)) +
  geom_boxplot(alpha = 0.5)

violin <- ggplot(pet_happy, aes(pet, happiness, fill=pet)) +
  geom_violin(alpha = 0.5)

ggsave("demog_violin_plot.png", width = 5, height = 7)

ggsave("demog_box_plot.jpg", plot = box, width = 5, height = 7)

```


## Combination Plots

### Violinbox plot

To demonstrate the use of `facet_grid()` for factorial designs, we create a new 
column called `agegroup` to split the data into participants older than the 
meadian age or younger than the median age. New factors will display in alphabetical order, so we can use the `factor()` function to set the levels in the order we want.

```{r violinbox, fig.cap="Violin-box plot"}

pet_happy %>%
  mutate(agegroup = ifelse(age<median(age), "Younger", "Older"),
         agegroup = factor(agegroup, levels = c("Younger", "Older"))) %>%
  ggplot(aes(pet, happiness, fill=pet)) +
    geom_violin(trim = FALSE, alpha=0.5, show.legend = FALSE) +
    geom_boxplot(width = 0.25, fill="white") +
    facet_grid(.~agegroup) +
    scale_fill_manual(values = c("orange", "green"))

```

```{block, type="info"}
Set the `show.legend` argument to `FALSE` to hide the legend. We do this here because the x-axis already labels the pet types.
```

### Violin-point-range plot

You can use `stat_summary()` to superimpose a point-range plot showning the mean ± 1 SD. You'll learn how to write your own functions in the lesson on [Iteration and Functions](#func).

```{r stat-summary, fig.cap="Point-range plot using stat_summary()"}

ggplot(pet_happy, aes(pet, happiness, fill=pet)) +
  geom_violin(
    trim = FALSE,
    alpha = 0.5
  ) +
  stat_summary(
    fun.y = mean,
    fun.ymax = function(x) {mean(x) + sd(x)},
    fun.ymin = function(x) {mean(x) - sd(x)},
    geom="pointrange"
  )

```

### Violin-jitter plot

If you don't have a lot of data points, it's good to represent them individually. 
You can use `geom_point` to do this, setting `position` to "jitter".

```{r violin-jitter, fig.cap="Violin-jitter plot"}

pet_happy %>%
  sample_n(50) %>%  # choose 50 random observations from the dataset
  ggplot(aes(pet, happiness, fill=pet)) +
  geom_violin(
    trim = FALSE,
    draw_quantiles = c(0.25, 0.5, 0.75), 
    alpha = 0.5
  ) + 
  geom_point(position = "jitter", alpha = 0.7, size = 3)

```

### Scatter-line graph

If your graph isn't too complicated, it's good to also show the individual data 
points behind the line.

```{r scatter_line, fig.cap="Scatter-line plot"}

ggplot(x_vs_y, aes(x, y)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method="lm")

```

### Grid of plots {#cowplot}

You can use the [ `cowplot`](https://cran.r-project.org/web/packages/cowplot/vignettes/introduction.html) package to easily make grids of different graphs. First, you have to assign each plot a name. Then you list all the plots as the first arguments of `plot_grid()` and provide a list of labels.

```{r cowplot, message = FALSE, fig.cap="Grid of plots using cowplot"}

library(cowplot)

my_hist <- ggplot(pet_happy, aes(happiness, fill=pet)) +
  geom_histogram(
    binwidth = 1, 
    alpha = 0.5, 
    position = "dodge", 
    show.legend = FALSE
  )

my_violin <- ggplot(pet_happy, aes(pet, happiness, fill=pet)) +
  geom_violin(
    trim = FALSE,
    draw_quantiles = c(0.5), 
    alpha = 0.5, 
    show.legend = FALSE
  )

my_box <- ggplot(pet_happy, aes(pet, happiness, fill=pet)) +
  geom_boxplot(alpha=0.5, show.legend = FALSE)

my_density <- ggplot(pet_happy, aes(happiness, fill=pet)) +
  geom_density(alpha=0.5, show.legend = FALSE)

my_bar <- pet_happy %>%
  group_by(pet) %>%
  summarise(
    mean = mean(happiness),
    sd = sd(happiness)
  ) %>%
  ggplot(aes(pet, mean, fill=pet)) +
    geom_bar(stat="identity", alpha = 0.5, show.legend = FALSE) +
    geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.25)

plot_grid(
  my_violin, 
  my_box, 
  my_density, 
  my_bar, 
  labels = c("A", "B", "C", "D")
)

```

```{block, type="info"}
{#theme}
Once you load the cowplot package, your ggplot default theme will change. You can get back the default ggplot theme with `+ theme_set(theme_grey())`.
```

## Overlapping Data {#overlap}

### Discrete Data 

You can deal with overlapping data points (very common if you're using Likert scales) by reducing the opacity of the points. You need to use trial and error to adjust these so they look right.

```{r overlap_alpha, fig.cap="Deal with overlapping data using transparency"}

ggplot(overlap, aes(x, y)) +
  geom_point(size = 5, alpha = .05) +
  geom_smooth(method="lm")

```

{#geom_count}
Or you can set the size of the dot proportional to the number of overlapping 
observations using `geom_count()`.

```{r overlap_size, fig.cap="Deal with overlapping data using geom_count()"}

overlap %>%
  ggplot(aes(x, y)) +
  geom_count(color = "#663399")

```

Alternatively, you can transform your data to create a count column and use the 
count to set the dot colour.

```{r overlap-colour, fig.cap="Deal with overlapping data using dot colour"}

overlap %>%
  group_by(x, y) %>%
  summarise(count = n()) %>%
  ggplot(aes(x, y, color=count)) +
  geom_point(size = 5) +
  scale_color_viridis()

```

### Colours {#viridis}

The [viridis package](https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html) 
changes the colour themes to be easier to read by people with colourblindness 
and to print better in greyscale. Use `scale_color_viridis()` to set the colour 
palette and `scale_fill_viridis()` to set the fill palette in ggplot. If you need 
discrete (as opposed to continuous) colours, use `scale_color_viridis(discrete=TRUE)` 
or `scale_fill_viridis(discrete=TRUE)` instead. 


```{block, type="info"}
The newest version of `ggplot2` v3.0.0 has viridis built in. It uses `scale_colour_viridis_c()` and `scale_fill_viridis_c()` for continuous variables and `scale_colour_viridis_d()` and `scale_fill_viridis_d()` for discrete variables.
```

### Continuous Data

Even if the variables are continuous, overplotting might obscure any relationships 
if you have lots of data.

```{r overplot-point, fig.cap="Overplotted data"}
overplot %>%
  ggplot(aes(x, y)) + 
  geom_point()
```

{#geom_density2d}
Use `geom_density2d()` to create a contour map.

```{r density2d, fig.cap="Contour map with geom_density2d()"}
overplot %>%
  ggplot(aes(x, y)) + 
  geom_density2d()
```

You can use `stat_density_2d(aes(fill = ..level..), geom = "polygon")` to create 
a heatmap-style density plot. 

```{r density2d-fill, fig.cap="Heatmap-density plot"}
overplot %>%
  ggplot(aes(x, y)) + 
  stat_density_2d(aes(fill = ..level..), geom = "polygon") +
  scale_fill_viridis()
```


{#geom_bin2d}
Use `geom_bin2d()` to create a rectangular heatmap of bin counts. Set the 
`binwidth` to the x and y dimensions to capture in each box.

```{r bin2d, fig.cap="Heatmap of bin counts"}
overplot %>%
  ggplot(aes(x, y)) + 
  geom_bin2d(binwidth = c(1,1))
```

{#geom_hex}
Use `geomhex()` to create a hexagonal heatmap of bin counts. Adjust the 
`binwidth`, `xlim()`, `ylim()` and/or the figure dimensions to make the hexagons 
more or less stretched.

```{r overplot-hex, fig.width=6, fig.height=5, fig.cap="Hexagonal heatmap of bin counts"}
overplot %>%
  ggplot(aes(x, y)) + 
  geom_hex(binwidth = c(0.25, 0.25))
```

## Heat map {#geom_tile}

I've included the code for creating a correlation matrix from a table of variables, 
but you don't need to understand how this is done yet. We'll cover `mutate` and 
`gather` functions in the `dplyr` and `tidyr` lessons.

```{r}
# generate two sets of correlated variables (a and b)
heatmap <- tibble(
  a1 = rnorm(100),
  b1 = rnorm(100)
) %>% 
mutate(
  a2 = a1 + rnorm(100),
  a3 = a1 + rnorm(100),
  a4 = a1 + rnorm(100),
  b2 = b1 + rnorm(100),
  b3 = b1 + rnorm(100),
  b4 = b1 + rnorm(100)
) %>%
cor() %>% # create the correlation matrix
as.data.frame() %>% # make it a data frame
rownames_to_column(var = "V1") %>% # set rownames as V1
gather("V2", "r", a1:b4) # wide to long (V2)
```

Once you have a correlation matrix in the correct (long) format, it's easy to 
make a heatmap using `geom_tile()`.

```{r heatmap, fig.cap="Heatmap using geom_tile()"}
ggplot(heatmap, aes(V1, V2, fill=r)) +
  geom_tile() +
  scale_fill_viridis()
```

```{block, type="info"}
The file type is set from the filename suffix, or by 
specifying the argument `device`, which can take the following values: 
"eps", "ps", "tex", "pdf", "jpeg", "tiff", "png", "bmp", "svg" or "wmf".
```

## Interactive Plots {#plotly}

You can use the `plotly` package to make interactive graphs. Just assign your 
ggplot to a variable and use the function `ggplotly()`.

```{r plotly, message = FALSE, fig.cap="Interactive graph using plotly"}

demog_plot <- ggplot(pet_happy, aes(pet, happiness, fill=pet)) +
  geom_point(position = position_jitter(width= 0.2, height = 0), size = 2)

ggplotly(demog_plot)

```

```{block, type="info"}
Hover over the data points above and click on the legend items.
```

## Quiz {#ggplot-quiz}

1. Generate a plot like this from the built-in dataset `iris`. Make sure to include the custom axis labels.

    ```{r quiz-boxplot, echo = FALSE, fig.cap = ""}
    ggplot(iris, aes(Species, Petal.Width, fill = Species)) +
      geom_boxplot(show.legend = FALSE) +
      xlab("Flower Species") +
      ylab("Petal Width (in cm)")
    ```
    
    ```{r quiz-boxplot-answer, eval = FALSE, webex.hide = TRUE}
    ggplot(iris, aes(Species, Petal.Width, fill = Species)) +
      geom_boxplot(show.legend = FALSE) +
      xlab("Flower Species") +
      ylab("Petal Width (in cm)")
    
    # there are many ways to do things, the code below is also correct
    ggplot(iris) +
      geom_boxplot(aes(Species, Petal.Width, fill = Species), show.legend = FALSE) +
      labs(x = "Flower Species",
           y = "Petal Width (in cm)")
    ```


2. You have just created a plot using the following code. How do you save it?
    ```{r quiz-save, eval = FALSE}
    ggplot(cars, aes(speed, dist)) + 
      geom_point() +
      geom_smooth(method = lm)
    ```
    
    ```{r quiz-save-answer, echo = FALSE, results = 'asis'}
    c(
      "ggsave()",
      "ggsave(\"figname\")",
      answer = "ggsave(\"figname.png\")",
      "ggsave(\"figname.png\", plot = cars)"
    ) %>%
      webex::mcq() %>% cat()
    ```

  
3. Debug the following code.
    ```{r quiz-debug, eval = FALSE}
    ggplot(iris) +
      geom_point(aes(Petal.Width, Petal.Length, colour = Species)) +
      geom_smooth(method = lm) +
      facet_grid(Species)
    ```
    
    ```{r quiz-debug-answer, webex.hide = TRUE, fig.cap = ""}
    ggplot(iris, aes(Petal.Width, Petal.Length, colour = Species)) +
      geom_point() +
      geom_smooth(method = lm) +
      facet_grid(~Species)
    ```


4. Generate a plot like this from the built-in dataset `ChickWeight`.  

    ```{r, echo = FALSE, fig.cap = ""}
    ggplot(ChickWeight, aes(weight, Time)) +
      geom_hex(binwidth = c(10, 1)) +
      scale_fill_viridis_c()
    ```
    
    ```{r, eval = FALSE, webex.hide = TRUE}
    ggplot(ChickWeight, aes(weight, Time)) +
      geom_hex(binwidth = c(10, 1)) +
      scale_fill_viridis_c()
    ```

    
5. Generate a plot like this from the built-in dataset `iris`.

    ```{r quiz-cowplot, echo = FALSE, fig.height = 10, fig.cap = ""}
    pw <- ggplot(iris, aes(Petal.Width, color = Species)) +
      geom_density() +
      xlab("Petal Width (in cm)")
    
    pl <- ggplot(iris, aes(Petal.Length, color = Species)) +
      geom_density() +
      xlab("Petal Length (in cm)") +
      coord_flip()
    
    pw_pl <- ggplot(iris, aes(Petal.Width, Petal.Length, color = Species)) +
      geom_point() +
      geom_smooth(method = lm) +
      xlab("Petal Width (in cm)") +
      ylab("Petal Length (in cm)")
    
    cowplot::plot_grid(
      pw, pl, pw_pl, 
      labels = c("A", "B", "C"),
      nrow = 3
    )
    ```
    
    ```{r quiz-cowplot-answer, eval = FALSE, webex.hide = TRUE}
    pw <- ggplot(iris, aes(Petal.Width, color = Species)) +
      geom_density() +
      xlab("Petal Width (in cm)")
    
    pl <- ggplot(iris, aes(Petal.Length, color = Species)) +
      geom_density() +
      xlab("Petal Length (in cm)") +
      coord_flip()
    
    pw_pl <- ggplot(iris, aes(Petal.Width, Petal.Length, color = Species)) +
      geom_point() +
      geom_smooth(method = lm) +
      xlab("Petal Width (in cm)") +
      ylab("Petal Length (in cm)")
    
    cowplot::plot_grid(
      pw, pl, pw_pl, 
      labels = c("A", "B", "C"),
      nrow = 3
    )
    ```

## Exercises

Download the [formative exercises](formative_exercises/03_ggplot_stub.Rmd). See the [answer demo](formative_exercises/03_ggplot_answers.html) to see what your plots should look like (this doesn't contain the answer code).


### Common Plots

Generate a violin plot, boxplot, histogram, density plot, and column plot for the following data. 

```{r}

# dog weights estimated from http://petobesityprevention.org/ideal-weight-ranges/

dogs <- tibble(
  breed = rep(c("beagle", "boxer", "bulldog"), each = 100),
  weight = c(
    rnorm(100, 24, 6),
    rnorm(100, 62.5, 12.5),
    rnorm(100, 45, 5)
  )
)

```

*Basic*: Create each plot.

*Intermediate*: Change the axis labels and colours. Save each plot as a PNG file.

*Advanced*: Create a grid of the first four plot styles (exclude the column plot). In your RMarkdown file, display just the graph, but not the `r` code for the graph.


### Two continuous variables

Represent the relationships among moral, sexual and pathogen disgust scores from the dataset [disgust_scores.csv](data/disgust_scores.csv). 

*Basic*: Graph the linear relationship between moral and pathogen disgust. Make sure the axes run from the minimum to maximum possible scores on both axes. Give the graph an appropriate title and axis lables.
  
  
*Intermediate*: Create a 2d density plot of the relationship between pathogen and sexual disgust. 

Use `stat_density_2d(aes(fill = ..level..), geom = "polygon", n = ?, 
h = c(?, ?))`, set n and h to values that make the graph look good, and 
figure out what `n` and `h` represent.

*Advanced*: Create a 3x3 grid of plots with columns representing the x-axis and 
rows representing the y-axis. Put a density plot of each variable along the 
diagonal. Make sure the graphs have appropriate titles and axis labels and 
that the range of the axes are the same in all graphs.

|              | moral   | sexual  | pathogen |
|--------------|---------|---------|----------|
| **moral**    | density | line    | line     | 
| **sexual**   | line    | density | line     |
| **pathogen** | line    | line    | density  |



### Many correlated variables

*Basic*: Create a heatmap of the relationships among all the questions in 
[disgust_cors.csv](data/disgust_cors.csv) 
(the correlations have already been calculated for you). 

*Intermediate*: Figure out how to rotate the text on the x-axis so it's readable.
