---
title: 'Formative Exercise 02: MSc Data Skills Course'
author: "Psychology, University of Glasgow"
output: html_document
---

```{r setup, include=FALSE}
# please do not alter this code chunk
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
set.seed(90210) # makes sure random numbers are reproducible
```

## Vectors

### Question 1

The built-in vector `letters` contains the letters of the English alphabet.  Use an indexing vector of integers to extract the letters that spell 'cat'.

```{r Q1}
cat <- NULL
```


### Question 2

The function `colors()` returns all of the color names that R is aware of. What is the length of the vector returned by this function? (Use code to find the answer.)

```{r Q2}
col_length <- NULL
```

### Question 3

The function call `runif(1000, 0, 1)` will draw 1000 numbers from a uniform distribution from 0 to 1, which simulates the p-values that you would get from 1000 experiments where the null hypothesis is true. Store the result of this call in `pvals`. Create a logical vector called `is_sig` that is `TRUE` if the corresponding element of `pvals` is less than .05, `FALSE` otherwise (hint: vectorized operations from the last lession), then use this logical vector to pull out those p-values. Finally, calculate the proportion of those p-values that were significant.

```{r Q3}
pvals <- runif(1000, 0, 1)

is_sig <- NULL

prop_sig <- NULL

```


## Tabular Data 

### Question 4

Create a tibble with the columns `name`, `age`, and `country` of origin for 3 people you know.

```{r Q4}

people <- NULL

```


### Question 5

Convert the built-in base R `iris` dataset to a tibble, and store it in the variable `iris2`.

```{r Q5}
iris2 <- NULL
```


### Question 6

Create a tibble that has the structure of the table below, using the minimum typing possible. (Hint: `rep()`).  Store it in the variable `my_tbl`.

ID |  A | B   | C
--|-----|-----|---
1	| A1	| B1	| C1	
2	| A1	| B2	| C1	
3	| A1	| B1	| C1	
4	| A1	| B2	| C1	
5	| A2	| B1	| C1	
6	| A2	| B2	| C1	
7	| A2	| B1	| C1	
8	| A2	| B2	| C1

```{r Q6}
my_tbl <- NULL  
```


## Data Import

### Question 7

Download the dataset [disgust_scores.csv](https://psyteachr.github.io/msc-data-skills/data/disgust_scores.csv) and read it into a table.

```{r Q7, message = FALSE}
disgust <- NULL
```


### Question 8

Override the default column specifications to skip the `id` column.

```{r Q8, message = FALSE}
disgust_skip <- NULL
```

### Question 9
  
How many rows and columns are in the dataset from question 7?

```{r Q9}
disgust_rows <- NULL
disgust_cols <- NULL
```


## Answer Checks

You've made it to the end. Make sure you are able to knit this document to HTML. You can check your answers below in the knit document.

```{r answer-checks, echo = FALSE, warning=FALSE, results='asis'}
# do not edit
Q <- c()
Q["1"] <- all.equal(cat, c("c", "a", "t"))
Q["2"] <- all.equal(col_length, 657L)
Q["3"] <- all.equal(prop_sig, 0.046)
Q["4"] <- c(all.equal(names(people), c("name", "age", "country")),
          nrow(people) == 3L) %>% all()
Q["5"] <- all.equal(iris, iris2, check.attributes = FALSE)
Q["6"] <- c(all.equal(names(my_tbl), c("ID", "A", "B", "C")),
          nrow(my_tbl) == 8L) %>% all()
Q["7"] <- c(all.equal(names(disgust), c("id", "user_id", "date", "moral", "pathogen", "sexual")), 
          nrow(disgust) == 20000L) %>% all()
Q["8"] <- c(all.equal(names(disgust_skip), c("user_id", "date", "moral", "pathogen", "sexual")), 
          nrow(disgust_skip) == 20000L) %>% all()
Q["9"] <- c(all.equal(disgust_rows, 20000L),
          all.equal(disgust_cols, 6)) %>% all()

ans <- sapply(Q, isTRUE)

knitr::kable(data.frame(
  Question = paste0("<a href='#question-", names(Q), "'>Question ", names(Q), "</a>"),
  Answer = ifelse(ans, "correct", "incorrect")
))
```
