---
title: "General Linear Models"
author: "Data Skills Teaching Team"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Add-on packages

Did you use any add-on packages when generating your answers?  If so, load them in the chunk below.

```{r add_ons}
library(tidyverse)
```

# Tasks 1-4: The `iris` dataset

There is a built-in dataset called `iris` that has measurements of different parts of flowers. (See `?iris` for information about the dataset.)


## Task 1

Use ggplot2 to make a scatterplot that visualizes the relationship between sepal length (horizontal axis) and petal width (vertical axis).

```{r T1}
ggplot(iris, aes(Sepal.Length, Petal.Width)) +
  geom_jitter(alpha = .2) # overplotting! watch out!
```


## Task 2

Run a regression model that predicts the petal width from the sepal length, and store the model object in the variable `iris_mod`.  End the block by printing out the summary of the model.

```{r T2}
iris_mod <- lm(Petal.Width ~ Sepal.Length, iris)

# TODO: print out the model summary
summary(iris_mod)
```


## Task 3

Make a histogram of the residuals of the model using ggplot2. (HINT: you may need to put the residuals into a table)

```{r T3}
ggplot(tibble(residual = residuals(iris_mod)), aes(residual)) +
  geom_histogram()
```


## Task 4

Write code to predict the petal width for two plants, the first with a sepal length of 5.25cm, and the second with a sepal length of 6.3cm.  Store the vector of predictions in the variable `iris_pred`.

```{r T4}
iris_pred <- predict(iris_mod, tibble(Sepal.Length = c(5.25, 6.3)))

# print them
iris_pred
```


# Simulating data from the linear model

## Task 5

```{r task5_setup, include = FALSE}
# DO NOT ALTER, REMOVE, OR MOVE THIS BLOCK.
set.seed(992) # for reproducibility
```

*NOTE: You can knit this file to html to see formatted versions of the equations below (which are enclosed in `$` characters); alternatively, if you find it easier, you can hover your mouse pointer over the `$` in the code equations to see the formatted versions.* 

Write code to randomly generate 10 Y values from a simple linear regression model with an intercept of 3 and a slope of -7.  Recall the form of the linear model:

$Y_i = \beta_0 + \beta_1 X_i + e_i$

The residuals ($e_i$s) are drawn from a normal distribution with mean 0 and variance $\sigma^2 = 4$, and $X$ is the vector of integer values from 1 to 10. Store the 10 observations in the variable `Yi` below. (NOTE: the standard deviation is the *square root* of the variance, i.e. $\sigma$; `rnorm()` takes the standard deviation, not the variance, as its third argument).

```{r T5}
x <- 1:10
err <- rnorm(10, sd = 2)
Yi <- 3 - 7 * x + err

# print Yi
Yi
```

