---
title: 'Formative Exercise 01: MSc Data Skills Course'
author: "Psychology, University of Glasgow"
output: html_document
---

```{r setup, include=FALSE}
# please do not alter this code chunk
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

Create a variable called `my_pals` containing the names of three of your friends. This should be a three-element vector of type `character`.

```{r Q1}
my_pals <- NULL
```

## Question 2

The chunk below creates an integer vector `ages` with the ages of five people. (Note: defining a number with the sufficx `L`, e.g., `26L`, defines `26` as type integer rather than type double.)  Calculate how old they will be in ten years using a vectorized operation and store the result in the variable `ages_10`.

```{r Q2}
ages <- c(26L, 13L, 47L, 62L, 18L)
```

## Question 3

The code below has an error and won't run, so evaluation of the block has been turned off (`eval = FALSE`). Fix the code and once you've fixed it, turn evaluation back on again (`eval = TRUE`).

```{r Q3, eval = FALSE}
# I wan't to use the cowsay package to make a cat
# say meow. Why won't this work? What am I missing?

say("meow")
```

## Question 4

Call the `rnorm()` function to generate 10 random values from a normal distribution with a mean of 800 and a standard deviation of 20, and store the result in the variable `random_vals`.

```{r Q4}
random_vals <- NULL
```

## Question 5

Write code that creates a logical vector `is_sig` that represents whether each of the p values in the vector `pvals` is less than .05. (Hint: type `1:11 < 7` in the console and see what you get).

```{r Q5}
pvals <- c(.2, .001, .78, .04, .06, 10e-9)

is_sig <- NULL # something with pvals
```

## All done!

You've made it to the end. Make sure you are able to compile this document to HTML. Upload the RMarkdown file to the Moodle site.