---
title: 'Formative Exercise 05: MSc Data Skills Course'
author: "Psychology, University of Glasgow"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library("ukbabynames")
library("lubridate")
library("tidyverse")

## convert to a tibble
ukb <- as_tibble(ukbabynames)
```

### Easy

1. How many records are in the dataset?

```{r}
count(ukb) ## or: nrow(ukb)
```

2. Remove the column `rank` from the dataset.

```{r}
ukb %>%
  select(-rank)
```

3. What is the range of birth years contained in the dataset?

```{r}
ukb %>%
  summarise(minyear = min(year),
            maxyear = max(year))
```

4. Pull out the babies named Hermione.

```{r}
ukb %>%
  filter(name == "Hermione")
```

5. Sort the dataset by sex and then by year (descending) and then by rank (descending).

```{r}
ukb %>%
  arrange(sex, desc(year), desc(rank))
```

6. Create a new column, `decade`, that contains the decade of birth (1990, 2000, 2010).  Hint: see `?floor`

```{r}
ukb %>%
  mutate(decade = floor(year / 10) * 10)
```

7. Pull out the male babies named Courtney that were born between 1998 and 2001 (inclusive).

```{r}
ukb %>%
  filter(name == "Courtney", sex == "M",
         year >= 1998, year <= 2001)
```


8. How many distinct names are represented in the dataset?

```{r}
## how many distinct names are represented in the dataset?
ukb %>%
  distinct(name) %>%
  count()
```

9. Pull out all the female babies named Frankie that were born before 1999 or after 2010

```{r ex1}
ukb %>%
  filter(name == "Frankie", sex == "F",
         (year < 1999) | (year > 2010))
```

10. How many total babies in the dataset were named 'Emily'?

```{r ex2}
## 
ukb %>%
  filter(name == "Emily") %>%
  summarise(tot = sum(n)) %>%
  pull(tot)
```

11. How many distinct names are there for each sex?

```{r ex3}
ukb %>% 
  group_by(sex) %>%
  distinct(name) %>%
  count()
```

12. What is the most popular name in the dataset?

```{r ex4}
ukb %>%
  group_by(name) %>%
  summarise(tot = sum(n)) %>%
  arrange(desc(tot)) %>%
  slice(1) %>%
  pull(name)
```

13. How many babies were born each year for each sex?  Make a plot.

```{r ex5}
babes_per_year <- ukb %>%
  group_by(year, sex) %>%
  summarise(tot = sum(n))

ggplot(babes_per_year, aes(year, tot, color = sex)) +
  geom_line()
```

