---
title: 'Formative Exercise 07: MSc Data Skills Course'
author: "Psychology, University of Glasgow"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library("tidyverse")
```


1. Set the vector `v1` equal to the following: "11, 13, 15, 17, 19, ..., 101 (use a function; don't just type all the numbers).

```{r Q1}

v1 <- seq(11, 101, by = 2)

```

2. Set the vector `v2` equal to the following: "A" "A" "B" "B" "C" "C" "D" "D" "E" "E" (note the letters are all uppercase).

```{r Q2}

v2 <- rep(LETTERS[1:5], each = 2)

```

3. Set the vector `v3` equal to the following: "a" "b" "c" "d" "e" "a" "b" "c" "d" "e" (note the letters are all lowercase)

```{r Q3}

v3 <- rep(letters[1:5], 2)

```

4. Set the vector `v4` equal to the words "dog" 10 times, "cat" 9 times, "fish" 6 times, and "ferret" 1 time. 

```{r}

pets <- c("dog", "cat", "fish", "ferret")
pet_n <- c(10, 9, 6, 1)
v4 <- rep(pets, times = pet_n)
  
```

5. Write a function called `my_add` that adds two numbers (`x` and `y`) together and returns the results.

```{r Q5}
my_add <- function(x, y) {
  x+y
}
```

6. Copy the function `my_add` below and add an error message if `x` or `y` are not numbers.

```{r Q6}
my_add <- function(x, y) {
  if (!is.numeric(x) | !is.numeric(y)) stop("x and y must be numbers")
  x+y
}
```

7. Create a tibble called `dat` that contains 20 rows and three columns: `id` (integers 101 through 120), `pre` and `post` (both 20-item vectors of random numbers from a normal distribution with mean = 0 and sd = 1).

```{r Q7}

dat <- tibble(
  id = 101:120,
  pre = rnorm(20),
  post = rnorm(20, 1)
)

```

8. Run a *paired-samples* t-test comparing `pre` and `post`. (check the help for `t.test`)

```{r Q8}

t.test(dat$post, dat$pre, paired = TRUE)

```

9. Use `broom::tidy` to save the results of the t-test in question 8 in a table called `stats`. (Hint: you can use `knitr::kable(stats)` to display the table in your knitted version.

```{r Q9}
stats <- t.test(dat$post, dat$pre, paired = TRUE) %>%
  broom::tidy()

knitr::kable(stats)
```

10. Create a function called `report_t` that takes a data table as an argument and returns the result of a paired-samples t-test between the columns `pre` and `post` in the following format: "The mean increase from pre-test to post-test was #.###: t(#) = #.###, p = 0.###, 95% CI = [#.###, #.###]." Hint: look at the function `paste0` (simpler) or `sprintf` (complicated but more powerful).

NB: Make sure all numbers are reported to three decimal places (except degrees of freedom).

```{r Q10}

report_t <- function(data) {
  stats <- t.test(data$post, data$pre, paired = TRUE) %>%
    broom::tidy()
  
  diff <- pull(stats, estimate) %>% round(3) 
  t <- pull(stats, statistic) %>% round(3)
  p <- pull(stats, p.value) %>% round(3)
  df <- pull(stats, parameter)
  ci1 <- pull(stats, conf.low) %>% round(3)
  ci2 <- pull(stats, conf.high) %>% round(3)
  
  paste0("The mean increase from pre-test to post-test was ", diff, 
         ": t(", df, ") = ", t, 
         ", p = ", p, 
         ", 95% CI = [", ci1, ", ", ci2, "].")
}

```

```{r Q10b}
# sprintf() is a complicated function, but can be easier to use in long text strings with a lot of things to replace

report_t <- function(data) {
  stats <- t.test(data$post, data$pre, paired = TRUE) %>%
    broom::tidy()

  sprintf("The mean increase from pre-test to post-test was %.3f: t(%.0f) = %.3f, p = %.3f, CI = [%.3f, %.3f].",
          pull(stats, estimate), 
          pull(stats, parameter),
          pull(stats, statistic),
          pull(stats, p.value),
          pull(stats, conf.low),
          pull(stats, conf.high)
  )
}

```


`r report_t(dat)`


