---
title: 'Formative Exercise 07: MSc Data Skills Course'
author: "Psychology, University of Glasgow"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library("tidyverse")
```


1. Set the vector `v1` equal to the following: "11, 13, 15, 17, 19, ..., 101 (use a function; don't just type all the numbers).

```{r Q1}

v1 <- NULL

```

2. Set the vector `v2` equal to the following: "A" "A" "B" "B" "C" "C" "D" "D" "E" "E" (note the letters are all uppercase).

```{r Q2}

v2 <- NULL

```

3. Set the vector `v3` equal to the following: "a" "b" "c" "d" "e" "a" "b" "c" "d" "e" (note the letters are all lowercase)

```{r Q3}

v3 <- NULL

```

4. Set the vector `v4` equal to the words "dog" 10 times, "cat" 9 times, "fish" 6 times, and "ferret" 1 time. 

```{r}

v4 <- NULL
  
```

5. Write a function called `my_add` that adds two numbers (`x` and `y`) together and returns the results.

```{r Q5}

my_add <- NULL

```

6. Copy the function `my_add` below and add an error message if `x` or `y` are not numbers.

```{r Q6}

my_add <- NULL

```

7. Create a tibble called `dat` that contains 20 rows and three columns: `id` (integers 101 through 120), `pre` and `post` (both 20-item vectors of random numbers from a normal distribution with mean = 0 and sd = 1).

```{r Q7}

dat <- NULL

```

8. Run a *paired-samples* t-test comparing `pre` and `post`. (check the help for `t.test`)

```{r Q8}

# paired-samples t-test

```

9. Use `broom::tidy` to save the results of the t-test in question 8 in a table called `stats`. (Hint: you can use `knitr::kable(stats)` to display the table in your knitted version.

```{r Q9}

stats <- NULL

```

10. Create a function called `report_t` that takes a data table as an argument and returns the result of a paired-samples t-test between the columns `pre` and `post` in the following format: "The mean increase from pre-test to post-test was #.###: t(#) = #.###, p = 0.###, 95% CI = [#.###, #.###]." Hint: look at the function `paste0` (simpler) or `sprintf` (complicated but more powerful).

NB: Make sure all numbers are reported to three decimal places (except degrees of freedom).

```{r Q10}

report_t <- NULL

```

