---
title: 'Exercise 01: MSc Data Skills Course'
author: "Psychology, University of Glasgow"
output: html_document
---

```{r setup, include=FALSE}
# please do not alter this code chunk
knitr::opts_chunk$set(echo = TRUE, message = FALSE, error = TRUE)
```

Edit the code chunks below and knit the document.

## Creating and manipulating variables

Create objects called `a` and `b` containing the numbers 10 and 12. Use the addition operator (`+`) to add them and assign the results of that to an object called `c`.

```{r Q1a}
a <- 10
b <- 12
c <- a + b
```


Create objects called `first_name` and `last_name` containing your first and last names. Use the `paste()` function to combine them and assign the results of that to an object called `full_name`.

```{r Q1b}
first_name <- "Lisa"
last_name  <- "DeBruine"
full_name  <- paste(first_name, last_name)
```


## Vectors

Variables can be more than one value. This is called a "vector". Use the function `c()` to group values into a vector. 

The chunk below creates a vector called `ages` with the ages of five people. Use functions and operations on `ages` to answer the following questions:

* `ages_10`: How old will each person be in 10 years?
* `mean_ages`: What is the mean age of the 5 people?
* `sd_ages`: What is the standard deviation of `ages`?
* `n_ages`: How many ages are listed in the vector `ages`?

```{r Q2a}
ages <- c(26, 13, 47, 62, 18)

ages_10   <- ages + 10
mean_ages <- mean(ages)
sd_ages   <- sd(ages)
n_ages    <- length(ages)
```


Replace {?} below with inline R to describe the sample from `ages2`. Knit to see the result.

There are `r n_ages` people with a mean age of `r mean_ages` years (SD = `r sd_ages`).


## Installing Packages

Install the CRAN package called "fortunes". Run the code to do and include it in the code chunk below, but comment it out. It is bad practice to write a script that installs a package without the user having the option to cancel. Also, some packages take a long time to load, so you won't want to install them every time you run a script.

```{r Q3a}
# comment out the installation code
# install.packages("fortunes")
```


The code below has an error and won't run. Fix the code.

```{r Q3b}
# alternatively, you could load the library first: library(fortunes)

fortunes::fortune()
```


## Functions and Help

Call the `rnorm()` function to generate 10 random values from a normal distribution with a mean of 800 and a standard deviation of 20, and store the resulting vector in the object `random_vals`.

```{r Q4a}
random_vals <- rnorm(10, 800, 20)
```


Use the help function to figure out what argument you need to set to ignore `NA` values when calculating the mean of `the_values`. Change the function below to store the mean of `the_values` in the variable `the_mean`.

```{r Q4b}
the_values <- c(1,1,1,2,3,4,6,8,9,9, NA) # do not alter this line
the_mean   <- mean(the_values, na.rm = TRUE)
```


Figure out what the function `seq()` does. Use the function to set `tens` to the vector `c(0, 10, 20, 30, 40, 50 ,60, 70 ,80 ,90, 100)`. Set `bins6` to the cutoffs if you wanted to divide the numbers 0 to 100 into 6 bins. For example, dividing 0 to 100 into 4 bins results in the cutoffs `c(0, 25, 50, 75, 100)`,

```{r Q4c}
tens  <- seq(from = 0, to = 100, by = 10)
bins6 <- seq(0, 100, length.out = 7)
```


Figure out how to use the `paste()` function to paste together strings with forward slashes ("/") instead of spaces. Use `paste()` to set `my_dir` to "my/project/directory".

```{r Q4d}
my_dir <- paste("my", "project", "directory", sep = "/")
```
