---
title: 'Formative Exercise 02: MSc Data Skills Course'
author: "Psychology, University of Glasgow"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
set.seed(90210) # makes sure random numbers are reproducible
```

## Vectors

### Question 1

The built-in vector `letters` contains the letters of the English alphabet.  Use an indexing vector of integers to extract the letters that spell 'cat'.

```{r Q1}
cat <- letters[c(3, 1, 20)]
```


### Question 2

The function `colors()` returns all of the color names that R is aware of. What is the length of the vector returned by this function? (Use code to find the answer.)

```{r Q2}
col_length <- length(colors())
```

### Question 3

The function call `runif(1000, 0, 1)` will draw 1000 numbers from a uniform distribution from 0 to 1, which simulates the p-values that you would get from 1000 experiments where the null hypothesis is true. Store the result of this call in `pvals`. Create a logical vector called `is_sig` that is `TRUE` if the corresponding element of `pvals` is less than .05, `FALSE` otherwise (hint: vectorized operations from the last lession), then use this logical vector to pull out those p-values. Finally, calculate the proportion of those p-values that were significant.

```{r Q3}
pvals <- runif(1000, 0, 1)

is_sig <- pvals < .05

prop_sig <- length(pvals[is_sig]) / length(pvals)

# alternatively:
prop_sig <- mean(is_sig)
prop_sig <- mean(pvals < .05)

```


## Tabular Data 

### Question 4

Create a tibble with the columns `name`, `age`, and `country` of origin for 3 people you know.

```{r Q4}

# you can do this with the tibble function
people <- tibble(name = c("Lisa", "Ben", "Robbie"),
                 age = c(42, 43, 11),
                 country = c("US", "UK", "UK") )

# also note:
# you can type this in row by row, rather than column by column,
# using the 'tribble' function
people <- tribble(~name,  ~age, ~country,
                  "Lisa",   42,  "US",
                  "Ben",    43,  "UK",
                  "Robbie", 11,  "UK")

```


### Question 5

Convert the built-in base R `iris` dataset to a tibble, and store it in the variable `iris2`.

```{r Q5}
iris2 <- as_tibble(iris)
```


### Question 6

Create a tibble that has the structure of the table below, using the minimum typing possible. (Hint: `rep()`).  Store it in the variable `my_tbl`.

ID |  A | B   | C
--|-----|-----|---
1	| A1	| B1	| C1	
2	| A1	| B2	| C1	
3	| A1	| B1	| C1	
4	| A1	| B2	| C1	
5	| A2	| B1	| C1	
6	| A2	| B2	| C1	
7	| A2	| B1	| C1	
8	| A2	| B2	| C1

```{r Q6}
my_tbl <- tibble(ID = 1:8,
                 A = rep(c("A1", "A2"), each = 4),
                 B = rep(c("B1", "B2"), 4),
                 C = "C1")    
```


## Data Import

### Question 7

Download the dataset [disgust_scores.csv](https://psyteachr.github.io/msc-data-skills/data/disgust_scores.csv) and read it into a table.

```{r Q7, message = FALSE}
# change to the location to where you put your csv file
disgust <- read_csv("https://psyteachr.github.io/msc-data-skills/data/disgust_scores.csv")
```


### Question 8

Override the default column specifications to skip the `id` column.

```{r Q8, message = FALSE}
my_cols <- cols(
  id = col_skip()
)

disgust_skip <- read_csv("https://psyteachr.github.io/msc-data-skills/data/disgust_scores.csv", col_types = my_cols)
```

### Question 9
  
How many rows and columns are in the dataset from question 7?

```{r Q9}

## dim() returns a vector c(rows, cols)
dimensions <- dim(disgust)
disgust_rows <- dimensions[1]
disgust_cols <- dimensions[2]

## nrow() returns the number of rows
disgust_rows <- nrow(disgust)

## ncol() returns the number of columns
disgust_cols <- ncol(disgust)
```


## Answer Checks

You've made it to the end. Make sure you are able to knit this document to HTML. You can check your answers below in the knit document.

```{r answer-checks, echo = FALSE, warning=FALSE, results='asis'}
# do not edit
Q <- c()
Q[1] <- all.equal(cat, c("c", "a", "t"))
Q[2] <- all.equal(col_length, 657L)
Q[3] <- all.equal(prop_sig, 0.046)
Q[4] <- c(all.equal(names(people), c("name", "age", "country")),
          nrow(people) == 3L)
Q[5] <- all.equal(iris, iris2, check.attributes = FALSE)
Q[6] <- c(all.equal(names(my_tbl), c("ID", "A", "B", "C")),
          nrow(my_tbl) == 8L)
Q[7] <- c(all.equal(names(disgust), c("id", "user_id", "date", "moral", "pathogen", "sexual")), 
          nrow(disgust) == 20000L)
Q[8] <- c(all.equal(names(disgust_skip), c("user_id", "date", "moral", "pathogen", "sexual")), 
          nrow(disgust_skip) == 20000L)
Q[9] <- c(all.equal(disgust_rows, 20000L),
          all.equal(disgust_cols, 6))

ans <- lapply(Q, function(x) mean(x==TRUE))

knitr::kable(data.frame(
  Question = paste0("<a href='#question-", 1:9, "'>Question ", 1:9, "</a>"),
  Answer = ifelse(ans==1, "correct", "incorrect")
))
```