---
title: 'Formative Exercise 02: MSc Data Skills Course'
author: "Psychology, University of Glasgow"
output: html_document
---

```{r setup, include=FALSE}
# please do not alter this code chunk
knitr::opts_chunk$set(echo = TRUE, message = FALSE, error = TRUE)
library(tidyverse)
```

## Load built-in datasets

List the datasets in `dplyr`.

```{r Q1}
# put code here

```


Load the built-in dataset `starwars`.

```{r Q2}
# put code here

```


Convert the built-in base R `mtcars` dataset to a tibble, and store it in the object `mt`.

```{r Q3}
mt <- NULL
```


## Import data from CSV and Excel files

Using the data directory created by `dataskills::getdata()` (or download [the zip file](https://psyteachr.github.io/msc-data-skills/data/data.zip), read "disgust_scores.csv" into a table.

```{r Q4}
disgust <- NULL
```

Override the default column specifications to skip the `id` column.

```{r Q5}
disgust_skip <- NULL
```


How many rows and columns are in the `disgust` dataset?

```{r Q6}
disgust_rows <- NULL
disgust_cols <- NULL
```


Load the data in "data/stroop.csv" as `stroop1` and "data/stroop.xlsx" as `stroop2`.

```{r Q7}
stroop1 <- NULL
stroop2 <- NULL
```

Use `glimpse()` to figure out the difference between the two data tables and fix the problem.

```{r Q8}
stroop2b <- NULL
```


## Create a data table

Create a tibble with the columns `name`, `age`, and `country` of origin for 2 people you know.

```{r Q9}
people <- NULL
```

Create a tibble that has the structure of the table below, using the minimum typing possible. (Hint: `rep()`).  Store it in the variable `my_tbl`.

ID |  A | B   | C
--|-----|-----|---
1	| A1	| B1	| C1	
2	| A1	| B2	| C1	
3	| A1	| B1	| C1	
4	| A1	| B2	| C1	
5	| A2	| B1	| C1	
6	| A2	| B2	| C1	
7	| A2	| B1	| C1	
8	| A2	| B2	| C1

```{r Q10}
my_tbl <- NULL  
```

## Understand the use the basic data types

Set the following objects to the number 1 with the indicated data type:

* `one_int` (integer)
* `one_dbl` (double)
* `one_chr` (character)

```{r Q11}
one_int <- NULL
one_dbl <- NULL
one_chr <- NULL
```

Set the objects `T_log`, `T_chr`, `T_int` and `T_dbl` to logical, character, integer and double values that will all be equal to `TRUE`. 

```{r Q12}
T_log <- NULL
T_chr <- NULL
T_int <- NULL
T_dbl <- NULL
```

Check your answers with this code:

```{r Q12-check}
# these should all evaluate to TRUE
c(
  T_log == TRUE,
  T_chr == TRUE,
  T_int == TRUE,
  T_dbl == TRUE,
  is.logical(T_log),
  is.character(T_chr),
  is.integer(T_int),
  is.double(T_dbl)
)
```


## Understand and use the basic container types

Create a vector of the numbers 3, 6, and 9.

```{r Q13}
threes <- NULL
```


The built-in vector `letters` contains the letters of the English alphabet.  Use an indexing vector of integers to extract the letters that spell 'cat'.

```{r Q14}
cat <- NULL
```


The function `colors()` returns all of the color names that R is aware of. What is the length of the vector returned by this function? (Use code to find the answer.)

```{r Q15}
col_length <- NULL
```


Create a named list called `col_types` where the name is each column in the built-in dataset `table1` and the value is the column data type (e.g., "double", "character", "integer", "logical").

```{r Q16}
col_types <- NULL
```


## Use vectorized operations

Set the object `x` to the integers 1 to 100. Use vectorised operations to set `y` to `x` squared. Use `plot(x, y)` to visualise the relationship between these two numbers.

```{r Q17}
x <- NULL
y <- NULL
```

Set `t` to the numbers 0 to 100 in increments of 0.1. Set `x` to the sine of `t` and `y` to the cosine of `t`. Plot `x` against `y`.

```{r Q18}
t <- NULL
x <- NULL
y <- NULL
```


The function call `runif(n, min, max)` will draw `n` numbers from a uniform distribution from `min` to `max`. If you set `n` to 10000, `min` to 0 and `max` to 1, this simulates the p-values that you would get from 10000 experiments where the null hypothesis is true. Create the following objects:

* `pvals`: 10000 simulated p-values using `runif()`
* `is_sig`: a logical vector that is `TRUE` if the corresponding element of `pvals` is less than .05, `FALSE` otherwise
* `sig_vals`: a vector of just the significant p-values
* `prop_sig`: the proportion of those p-values that were significant

```{r Q19}
set.seed(8675309) # ensures you get the same random numbers each time you run this code chunk

pvals    <- NULL
is_sig   <- NULL
sig_vals <- NULL
prop_sig <- NULL
```

