---
title: 'Formative Exercise 07: MSc Data Skills Course'
author: "Psychology, University of Glasgow"
output: html_document
---

```{r setup, include=FALSE}
# please do not alter this code chunk
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library("tidyverse")
set.seed(1618034) # makes sure random numbers are reproducible
```

### Question 1

Set the vector `v1` equal to the following: 11, 13, 15, 17, 19, ..., 99, 101 (use a function; don't just type all the numbers).

```{r Q1}

v1 <- seq(11, 101, by = 2)

```

### Question 2

Set the vector `v2` equal to the following: "A" "A" "B" "B" "C" "C" "D" "D" "E" "E" (note the letters are all uppercase).

```{r Q2}

v2 <- rep(LETTERS[1:5], each = 2)

```

### Question 3

Set the vector `v3` equal to the following: "a" "b" "c" "d" "e" "a" "b" "c" "d" "e" (note the letters are all lowercase)

```{r Q3}

v3 <- rep(letters[1:5], 2)

```

### Question 4

Set the vector `v4` equal to the words "dog" 10 times, "cat" 9 times, "fish" 6 times, and "ferret" 1 time. 

```{r}

pets <- c("dog", "cat", "fish", "ferret")
pet_n <- c(10, 9, 6, 1)
v4 <- rep(pets, times = pet_n)
  
```

### Question 5

Write a function called `my_add` that adds two numbers (`x` and `y`) together and returns the results.

```{r Q5}
my_add <- function(x, y) {
  x+y
}
```

### Question 6

Copy the function `my_add` above and add an error message that returns "x and y must be numbers" if `x` or `y` are not both numbers.

```{r Q6}
my_add <- function(x, y) {
  if (!is.numeric(x) | !is.numeric(y)) stop("x and y must be numbers")
  x+y
}
```

### Question 7

Create a tibble called `dat` that contains 20 rows and three columns: `id` (integers 101 through 120), `pre` and `post` (both 20-item vectors of random numbers from a normal distribution with mean = 0 and sd = 1).

```{r Q7}

dat <- tibble(
  id = 101:120,
  pre = rnorm(20),
  post = rnorm(20)
)

```

### Question 8

Run a two-tailed, *paired-samples* t-test comparing `pre` and `post`. (check the help for `t.test`)

```{r Q8}

t <- t.test(dat$post, dat$pre, paired = TRUE)

t # prints results of t-test

```

### Question 9

Use `broom::tidy` to save the results of the t-test in question 8 in a table called `stats`.

```{r Q9, results='asis'}
stats <- t.test(dat$post, dat$pre, paired = TRUE) %>%
  broom::tidy()

knitr::kable(stats) # display the table
```

### Question 10

Create a function called `report_t` that takes a data table as an argument and returns the result of a two-tailed, paired-samples t-test between the columns `pre` and `post` in the following format: "The mean increase from pre-test to post-test was #.###: t(#) = #.###, p = 0.###, 95% CI = [#.###, #.###]." Hint: look at the function `paste0` (simpler) or `sprintf` (complicated but more powerful).

NB: Make sure all numbers are reported to three decimal places (except degrees of freedom).

```{r Q10}

report_t <- function(data) {
  stats <- t.test(data$post, data$pre, paired = TRUE) %>%
    broom::tidy()
  
  diff <- pull(stats, estimate) %>% round(3) 
  t <- pull(stats, statistic) %>% round(3)
  p <- pull(stats, p.value) %>% round(3)
  df <- pull(stats, parameter)
  ci1 <- pull(stats, conf.low) %>% round(3)
  ci2 <- pull(stats, conf.high) %>% round(3)
  
  paste0("The mean increase from pre-test to post-test was ", diff, 
         ": t(", df, ") = ", t, 
         ", p = ", p, 
         ", 95% CI = [", ci1, ", ", ci2, "].")
}

```

```{r Q10b}
# sprintf() is a complicated function, but can be easier to use in long text strings with a lot of things to replace

report_t <- function(data) {
  stats <- t.test(data$post, data$pre, paired = TRUE) %>%
    broom::tidy()

  sprintf("The mean increase from pre-test to post-test was %.3f: t(%.0f) = %.3f, p = %.3f, 95%% CI = [%.3f, %.3f].",
          pull(stats, estimate), 
          pull(stats, parameter),
          pull(stats, statistic),
          pull(stats, p.value),
          pull(stats, conf.low),
          pull(stats, conf.high)
  )
}

```


`r report_t(dat)`

## Answer Checks

You've made it to the end. Make sure you are able to knit this document to HTML. You can check your answers below in the knit document.

```{r answer-checks, echo = FALSE, warning=FALSE, results='asis'}
# do not edit
Q <- c()
Q["1"] <- c(all.equal(length(v1), 46),
            all.equal(v1[6], 21)) %>% all()
Q["2"] <- c(all.equal(length(v2), 10),
            all.equal(v2[6], "C")) %>% all()
Q["3"] <- c(all.equal(length(v3), 10),
            all.equal(v3[6], "a")) %>% all()
Q["4"] <- c(all.equal(length(v4), 26),
            all.equal(sum(v4 == "dog"), 10),
            all.equal(sum(v4 == "cat"), 9),
            all.equal(sum(v4 == "fish"), 6),
            all.equal(sum(v4 == "ferret"), 1)) %>% all()
Q["5"] <- ifelse(is.function(my_add),
            all.equal(my_add(10, -3), 7), FALSE)
Q["6"] <- if (is.function(my_add)) {
  c(all.equal(tryCatch(my_add("a", 1), error = function(e) {e$message}),"x and y must be numbers"), 
            all.equal(tryCatch(my_add(1.55, "a"), error = function(e) {e$message}), "x and y must be numbers")) } else { FALSE } %>% all()
Q["7"] <- c(is.data.frame(dat),
            all.equal(nrow(dat), 20),
            all.equal(sum(dat$id), 2210),
            names(dat) %in% c("id", "pre", "post")) %>% all()
Q["8"] <- c(is.list(t),
            all.equal(t$p.value, 0.2045149, tolerance = 1e-6)) %>% all()
Q["9"] <- c(is.data.frame(stats),
            all.equal(stats$p.value, 0.2045149, tolerance = 1e-6)) %>% all()
Q["10"] <- ifelse(is.function(report_t),
             all.equal(report_t(dat), "The mean increase from pre-test to post-test was 0.415: t(19) = 1.314, p = 0.205, 95% CI = [-0.246, 1.076]."), FALSE)
  
ans <- sapply(Q, isTRUE)

knitr::kable(data.frame(
  Question = paste0("<a href='#question-", names(Q), "'>Question ", names(Q), "</a>"),
  Answer = ifelse(ans, "correct", "incorrect")
))
```

