---
title: 'Formative Exercise 08: MSc Data Skills Course'
author: "Psychology, University of Glasgow"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library("tidyverse")
```

## Generate and compare random data

Generate 50 data points from a normal distribution with a mean of 0 
and SD of 1 (variable `a`).

```{r ex-1A}
a <- NULL
```

Generate another variable (`b`) that is equal to the sum of `a` and another 
50 data points from a normal distribution with a mean of 0.5 and SD of 1.

```{r ex-1B}
b <- NULL
```

Run a paired-samples t-test comparing `a` and `b`.

```{r ex-1C}
```

## Calculate Power

Calculate power for a two-tailed t-test with an alpha of .05 for detecting 
a difference between two independent samples of 50 with an effect size of 0.3.

Hint: You can use the `sim_t_ind` function from the [T-Test Class Notes](https://gupsych.github.io/data_skills/08_simulation.html#t-test).

```{r ex-2-sim}
power <- NULL

```

Compare this to the result of `power.t.test` for the same design.

```{r ex-2-function}
tpower <- NULL
```

Modify the `sim_t_ind` function to handle different 
sample sizes. Use it to calculate the power of the following design: 
20 observations from a normal distribution with mean of 10 and SD of 4 versus
30 observations from a normal distribution with a mean of 13 and an SD of 4.5.

```{r ex-3}
power2 <- NULL
```


Calculate power for a two-tailed t-test with an alpha of .005 for detecting 
a sex difference in height in a sample of 10 male and 10 female 20-year-olds? 
Get the data for the height of male and female 20-year-olds from the 
[US CDC Growth Chart Data Tables](https://www.cdc.gov/growthcharts/data/zscore/zstatage.csv).

```{r ex-4}
power3 <- NULL
```

## Poisson Distribution

The [poisson distribution(https://en.wikipedia.org/wiki/Poisson_distribution) 
is good for modeling the rate of something, like the number of texts you 
receive per day. Then you can test hypotheses like you receive more texts on 
weekends than weekdays. The poisson distribution gets more like a normal 
distribution when the rate gets higher, so it's most useful for low-rate events.

Use `ggplot` to create a histogram of 1000 random numbers from a poisson 
distribution with a `lambda` of 4. 

```{r ex-5}

```

## Binomial Distribution

Demonstrate to yourself that the binomial distribution looks like the normal 
distribution when the number of trials is greater than 10.

Hint: You can calculate the equivalent mean for the normal distribution as the 
number of trials times the probability of success (`binomial_mean <- trials * prob`) 
and the equivalent SD as the square root of the mean times one minus probability 
of success (`binomial_sd <- sqrt(binomial_mean * (1 - prob))`).

```{r ex-6}

```

## Correlation power simulation

Write a function to create a pair of variables of any size 
with any specified correlation. 

Hint: modify the code from the 
[Bivariate Normal](https://gupsych.github.io/data_skills/08_simulation.html#bvn) 
section from the class notes.

```{r ex-7A}
bvn2 <- NULL
```

Use `cor.test` to test the Pearson's product-moment 
correlation between two variables generated with your 
function having an `n` of 50 and a `rho` of 0.45.


```{r ex-7B1}

```

Make a new function that calculates power for `cor.test`. 
Remember, there are many, many ways to do things in R. 
The solution below is only one of many. The important thing is to 
create your function step-by-step, checking the accuracy at each step.

```{r ex-7C}
power.cor.test <- NULL
```

