---
title: "Iteration & Functions"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, error = TRUE)
library(tidyverse) ## contains purrr, tidyr, dplyr
# install.packages("broom")
library(broom) ## converts test output to tidy tables
```

## Resources

* [Iteration & Function](https://psyteachr.github.io/msc-data-skills/func.html) in Data Skills for Reproducible Science
* Chapters 19 and 21 of [R for Data Science](http://r4ds.had.co.nz)
* [RStudio Apply Functions Cheatsheet](https://github.com/rstudio/cheatsheets/raw/master/purrr.pdf)


## Iteration functions

### `rep()`

The function `rep()` lets you repeat the first argument a number of times.

Use `rep()` to create a vector of alternating `"A"` and `"B"` values of length 24.

```{r rep1}

```

If you don't specify what the second argument is, it defaults to `times`, repeating the vector in the first argument that many times. Make the same vector as above, setting the second argument explicitly.

```{r rep1-times}

```

If the second argument is a vector that is the same length as the first argument, each element in the first vector is repeated than many times. Use `rep()` to create a vector of 11 `"A"` values followed by 3 `"B"` values.

```{r rep-vector}

```

You can repeat each element of the vector a sepcified number of times using the `each` argument, Use `rep()` to create a vector of 12 `"A"` values followed by 12 `"B"` values.

```{r rep-each}

```

What do you think will happen if you set both `times` to 3 and `each` to 2?

```{r rep-times-each}

```


### `seq()`

The function `seq()` is useful for generating a sequence of numbers with some pattern.

Use `seq()` to create a vector of the integers 0 to 10.

```{r seq1-10}

```

You can set the `by` argument to count by numbers other than 1 (the default). Use `seq()` to create a vector of the numbers 0 to 100 by 10s.

```{r seq-by}

```

The argument `length.out` is useful if you know how many steps you want to divide something into. Use `seq()` to create a vector that starts with 0, ends with 100, and has 12 equally spaced steps (hint: how many numbers would be in a vector with 2 *steps*?).

```{r seq-length-out}

```

## Custom functions 

In addition to the built-in functions and functions you can access from packages, you can also write your own functions (and eventually even packages!).

### Structuring a function

The general structure of a function is as follows:

```{r}
function_name <- function(my_args) {
  # process the arguments
  # return some value
}
```

Here is a very simple function. Can you guess what it does?

```{r}
add1 <- function(my_number) {
  my_number + 1
}

add1(10)
```

Let's make a function that reports p-values in APA format (with "p = rounded value" when p >= .001 and "p < .001" when p < .001).

First, we have to name the function. You can name it anything, but try not to duplicate existing functions or you will overwrite them. For example, if you call your function `rep`, then you will need to use `base::rep()` to access the normal `rep` function. Let's call our p-value function `report_p` and set up the framework of the function.

```{r}
report_p <- NULL
```

### Arguments 

We need to add one *argument*, the p-value you want to report. The names you choose for the arguments are private to that argument, so it is not a problem if they conflict with other variables in your script. You put the arguments in the parentheses after `function` in the order you want them to default (just like the built-in functions you've used before). 

```{r}

```

### Argument defaults

You can add a default value to any argument. If that argument is skipped, then the function uses the default argument. It probably doesn't make sense to run this function without specifying the p-value, but we can add a second argument called `digits` that defaults to 3, so we can round p-values to 3 digits.

```{r}

```

Now we need to write some code inside the function to process the input arguments and turn them into a **return**ed output. Put the output as the last item in function. You might also see the returned output inside of the `return()` function. This does the same thing.

```{r}

```

When you run the code defining your function, it doesn't output anything, but makes a new object in the Environment tab under **`Functions`**. Now you can run the function.

```{r}
report_p(0.04869)
report_p(0.0000023)
```

### Scope

What happens in a function stays in a function. You can change the value of a variable passed to a function, but that won't change the value of the variable outside of the function, even if that variable has the same name as the one in the function.

```{r}
half <- function(x) {
  x <- x/2
  return(x)
}

x <- 10
list(
  "half(x)" = half(x),
  "x" = x
)
```


### Warnings and errors

What happens when you omit the argument for `p`? Or if you set `p` to 1.5 or "a"?

You might want to add a more specific warning and stop running the function code if someone enters a value that isn't a number. You can do this with the `stop()` function.

If someone enters a number that isn't possible for a p-value (0-1), you might want to warn them that this is probably not what they intended, but still continue with the function. You can do this with `warning()`.

```{r}

```

## Iterating your own functions

First, let's build up the code that we want to iterate.

### `rnorm()`

Create a vector of 20 random numbers drawn from a normal distribution with a mean of 5 and standard deviation of 1 using the `rnorm()` function and store them in the variable `A`.

```{r}
A <- NULL
```

### `tibble::tibble()`

A `tibble` is a type of table or `data.frame`. The function `tibble::tibble()` creates a tibble with a column for each argument. Each argument takes the form `column_name = data_vector`.

Create a table called `dat` including two vectors: `A` that is a vector of 20 random normally distributed numbers with a mean of 5 and SD of 1, and `B` that is a vector of 20 random normally distributed numbers with a mean of 5.5 and SD of 1.

```{r}
dat <- NULL
```

### `t.test`

You can run a Welch two-sample t-test by including the two samples you made as the first two arguments to the function `t.test`. You can reference one column of a table by its names using the format `table_name$column_name`

```{r}

```

You can also convert the table to long format using the `gather` function and specify the t-test using the format `dv_column~grouping_column`.

```{r}
longdat <- NULL

t.test(dv_column~grouping_column, data = longdat) 
```

### `broom::tidy()`

You can use the function `broom::tidy()` to extract the data from a statistical test in a table format. The example below pipes everything together and extracts a single value from this results table using `pull()`.

```{r}
tibble(
  A = rnorm(20, 5, 1),
  B = rnorm(20, 5.5, 1)
) %>%
  gather(group, score, A:B) %>%
  t.test(score~group, data = .) %>%
  broom::tidy() %>%
  pull(p.value)
```

### Turn into a function

First, name your function `t_sim` and wrap the code above in a function with no arguments. 

```{r}
t_sim <- function() {
  # add code above here
}
```

Run it a few times to see what happens.

```{r}
t_sim()
```

### `replicate()`

You can use the `replicate` function to run a function any number of times.

```{r}
replicate(3, rnorm(5))
```


Let's run the `t_sim` function 1000 times, assign the resulting p-values to a vector called `reps`, and check what proportion of p-values are lower than alpha (e.g., .05). This number is the power for this analysis.

```{r}

```

### Set seed

You can use the `set.seed` function before you run a function that uses random numbers to make sure that you get the same random data back each time. You can use any integer you like as the seed.

```{r set-seed-90210}
set.seed(90201)
```

Make sure you don't ever use `set.seed()` **inside** of a simulation function, or you will just simulate the exact same data over and over again.


### Add arguments

You can just edit your function each time you want to cacluate power for a different sample n, but it is more efficent to build this into your fuction as an arguments. Redefine `t_sim`, setting arguments for the mean and SD of group A, the mean and SD of group B, and the number of subjects per group. Give them all default values.


```{r}

```

Test your function with some different values to see if the results make sense.

```{r}

```

Use `replicate` to calculate power for 100 subjects/group with an effect size of 0.2 (e.g., A: m = 0, SD = 1; B: m = 0.2, SD = 1). Use 1000 replications.

```{r}

```

Compare this to power calculated from the `power.t.test` function.

```{r}
power.t.test(n = 100, delta = 0.2, sd = 1, type="two.sample")
```

Calculate power via simulation and `power.t.test` for the following tests:

* 20 subjects/group, A: m = 0, SD = 1; B: m = 0.2, SD = 1
* 40 subjects/group, A: m = 0, SD = 1; B: m = 0.2, SD = 1
* 20 subjects/group, A: m = 10, SD = 1; B: m = 12, SD = 1.5

```{r}

```