---
title: 'Formative Exercise 02: MSc Data Skills Course'
author: "Psychology, University of Glasgow"
date: "28 September 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Vectors

## Question 1

1. The built-in vector `letters` contains the letters of the English alphabet.  Use an indexing vector of integers to extract the letters that spell 'cat'.

```{r ex_vec_1, eval = FALSE}
letters[c(3, 1, 20)]
```

## Question 2

2. The function `colors()` returns all of the color names that R is aware of. What is the length of the vector returned by this function? (Use code to find the answer.)

```{r ex_vec_2, eval = FALSE}
length(colors())
```

## Question 3

3. The function call `runif(1000, 0, 1)` will draw 1000 numbers from a uniform distribution from 0 to 1, which simulates the p-values that you would get from 1000 experiments where the null hypothesis is true. Store the result of this call in `pvals`. Create a logical vector called `is_sig` that is `TRUE` if the corresponding element of `pvals` is less than .05, `FALSE` otherwise (hint: vectorized operations from the last lession), then use this logical vector to pull out those p-values. Finally, calculate the proportion of those p-values that were significant.

```{r ex_vec_3, eval = FALSE}
pvals <- runif(1000, 0, 1)

is_sig <- pvals < .05

sort(pvals[is_sig])

length(pvals[is_sig]) / length(pvals)
```

# Tabular Data 

## Question 1

1. Create a tibble with the name, age, and sex of 3-5 people whose names, ages, and sex you know.

```{r ex1-tab, eval = FALSE}

# or you can do this with the tibble function
family <- tibble(name = c("Lisa", "Ben", "Robbie"),
                 age = c(40, 41, 10),
                 sex = c("female", "male", "male") )

# also note:
# you can type this in row by row, rather than column by column,
# using the 'tribble' function
family <- tribble(~name,  ~age, ~sex,
                  "Lisa",   40,  "female",
                  "Ben",    41,  "male",
                  "Robbie", 10,  "male")

```

## Question 2

2. Convert the built-in base R `iris` dataset to a tibble, and store it in the variable `iris2`.

```{r ex_2_as_tib}
iris2 <- as_tibble(iris)
```

## Question 3

3. Create a tibble that has the structure of the table below, using the minimum typing possible. (Hint: `rep()`).  Store it in the variable `my_tbl`.

ID |  A | B   | C
--|-----|-----|---
1	| A1	| B1	| C1	
2	| A1	| B2	| C1	
3	| A1	| B1	| C1	
4	| A1	| B2	| C1	
5	| A2	| B1	| C1	
6	| A2	| B2	| C1	
7	| A2	| B1	| C1	
8	| A2	| B2	| C1

```{r ex_3_tib}
my_tbl <- tibble(ID = 1:8,
                 A = rep(c("A1", "A2"), each = 4),
                 B = rep(c("B1", "B2"), 4),
                 C = "C1")    
```

# Data Import

## Question 1

1. Download the dataset [disgust_scores.csv](data/disgust_scores.csv) and read it into a table.

```{r ex2-tab, message = FALSE}
# change the location to where you put your csv file
disgust <- read_csv("../data/disgust_scores.csv")
```

## Question 2

2. Override the default column specifications to skip the `id` column.

```{r ex3-tab, message = FALSE}
my_cols <- cols(
  id = col_skip()
)

disgust <- read_csv("../data/disgust_scores.csv", col_types = my_cols)
head(disgust)
```

## Question 3
  
3. How many rows and columns are in the dataset from question 3?

```{r ex4-tab}
## gives rows as "Observations"" and columns as "Variables"
glimpse(disgust)

## returns a list c(rows, cols)
dim(disgust)

## returns the number of rows
nrow(disgust)

## returns the number of columns
ncol(disgust)
```

# Pipes

## Question 1

1. Re-write the following sequence of commands into a single 'pipeline'.

```{r ex_pipe1}
# do not edit this chunk
x <- 1:20 # integers from 1:20
y <- rep(x, 2) # then repeat them twice
z <- sum(y) # and then take the sum
```

```{r ex_pipe1_ans}
x <- 1:20 %>% rep(2) %>% sum()
```

2. Deconstruct the pipeline below back into separate commands.

```{r ex_pipe2}
# do not edit this chunk
lager <- LETTERS[c(18, 5, 7, 1, 12)] %>%
  rev() %>%
  paste(collapse = "")
```

```{r ex_pipe2_ans}
regal <- LETTERS[c(18, 5, 7, 1, 12)]
reversed <- rev(regal)
lager <- paste(reversed, collapse = "") # make it into a string
```
