---
title: 'Formative Exercise 05: MSc Data Skills Course'
author: "Psychology, University of Glasgow"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library("ukbabynames")
library("lubridate")
library("tidyverse")

## convert to a tibble
ukb <- as_tibble(ukbabynames)
```

### Easy

1. How many records are in the dataset?

```{r}
```

2. Remove the column `rank` from the dataset.

```{r}
```

3. What is the range of birth years contained in the dataset?

```{r}
```

4. Pull out the babies named Hermione.

```{r}
```

5. Sort the dataset by sex and then by year (descending) and then by rank (descending).

```{r}
```

6. Create a new column, `decade`, that contains the decade of birth (1990, 2000, 2010).  Hint: see `?floor`

```{r}
```

7. Pull out the male babies named Courtney that were born between 1998 and 2001 (inclusive).

```{r}
```


8. How many distinct names are represented in the dataset?

```{r}
```

9. Pull out all the female babies named Frankie that were born before 1999 or after 2010

```{r ex1}
```

10. How many total babies in the dataset were named 'Emily'?

```{r ex2}
```

11. How many distinct names are there for each sex?

```{r ex3}
```

12. What is the most popular name in the dataset?

```{r ex4}
```

13. How many babies were born each year for each sex?  Make a plot.

```{r ex5}
```

